Composable Generalized Consensus

### 2016-01-05 paper/abstract3.tex
Some modifications to the abstract and beginning of the intro, doesn't compile anymore and I don't know why.

\abstract{
State-Machine Replication (SMR) is a basic building block of reliable
distributed systems. Improving SMR algorithms used by cloud-computing providers
can have a big impact on the reliability of the many computing services they
support. However, improving upon existing SMR algorithms is hard, often
justifying a publication in a top conference.

@@

We present a framework that enables building SMR algorithms out of independent modules. A module is an SMR algorithm that can abort its execution and pass the baton to a new module, allowing to build composite SMR algorithms which can dynamically switch between modules. 
Crucially, each module can be proved correct once and for all in isolation and our composition theorem guarantees that 
the composition of any number of correct modules is a correct SMR algorithm.

The key novelty of our framework is its practical applicability: we show that
a large class of SMR algorithms can be expressed in our framework by proposing
two specifications capturing the essence of modular SMR algorithms based on quorum systems. We
demonstrate that concrete algorithms can easily be obtained by refinement from
our abstract specifications, as evidenced by TLA+ specifications ready to be
checked by the TLC model checker.
}

>>>>

\abstract{
State-Machine Replication (SMR) is a basic building block of reliable
distributed systems. Improving SMR algorithms used by cloud-computing providers
can have a big impact on the reliability of the many computing services they
support. However, doing so is hard, often justifying a publication in a top conference.

@@

We present a framework that enables building SMR algorithms out of independent modules. 
A module is an SMR algorithm that can abort its execution and pass the baton to a new module, allowing to build composite SMR algorithms which can dynamically switch between modules. 
Crucially, each module can be analyzed once and for all in isolation; our composition theorem guarantees that 
the composition of any number of correct modules is a correct SMR algorithm.

The key novelty of our framework is its wide applicability: we show that
a large class of SMR algorithms can be expressed in our framework by proposing
two specifications capturing the essence of modular SMR algorithms based on quorum systems. We
demonstrate that concrete algorithms can easily be obtained by refinement from
our abstract specifications and we show that our framework enables formal specification and verification of composite SMR algorithms in TLA+ and Isabelle/HOL.
}

### 2016-01-05 paper/intro.tex
Some modifications to the abstract and beginning of the intro, doesn't compile anymore and I don't know why.

The State-Machine Replication technique (abbreviated SMR) allows building
reliable services on top of unreliable hardware. The basic idea is to
replicate a service over several servers, each executing the same sequence of
deterministic commands. In such a setup, any replica can be used to determine
the result of a command. Therefore, if a replica crashes, clients can be
transparently redirected to another one and will have the illusion of accessing
a centralized service at all times.

Making SMR algorithms efficient has been a long-standing challenge
and a multitude of algorithms have been proposed. Examples include
Paxos \cite{lamport2001paxos}, Fast Paxos \cite{Lamport06FastPaxos},
Disk Paxos \cite{GafniLamport03DiskPaxos}, Chain Replication
\cite{RenesseSchneider04ChainReplicationSupportingHighThroughputAvailability},
Ring Paxos \cite{MarandiETAL10RingPaxosHighthroughputAtomicBroadcastProtocol},
Egalitarian Paxos
\cite{MoraruAndersenKaminsky13ThereIsMoreConsensusEgalitarianParliaments},
Multi-Coordinated Paxos \cite{CamargosSchmidtPedone07MulticoordinatedPaxos},
Cheap Paxos \cite{LamportMassa04CheapPaxos}, Vertical Paxos
\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication}, Paxos-MIC
\cite{HurfinMoiseNarzul11AdaptiveFastPaxosMakingQuickEverlasting}, Mencius
\cite{MaoJunqueiraMarzullo08MenciusBuildingEfficientReplicatedStateMachine},
Fast Mencius \cite{WeiETAL13FastMenciusMenciusLowCommitLatency},
the Chandra-Toueg algorithm
\cite{ChandraToueg96UnreliableFailureDetectorsReliableDistributedSystems}, the
Ben-Or algorithm \cite{BenOr83AnotherAdvantageFreeChoiceCompletelyAsynchronous},
and Raft \cite{OngaroOusterhout14SearchUnderstandableConsensusAlgorithm}.

With the wide use of cloud computing, improvements in the performance and
efficiency of SMR algorithms have the potential to impact most online services:
advanced SMR algorithms can be implemented by cloud computing service providers
and benefit all their customer's applications.

However, devising new SMR algorithms, optimized with
particular situations or metrics in mind, is hard: today,
more than two decades after the first SMR algorithms
\cite{Lamport98ParttimeParliament,BirmanJoseph87ReliableCommunicationPresenceFailures,OkiLiskov88ViewstampedReplicationGeneralPrimaryCopy,DworkLynchStockmeyer84ConsensusPresencePartialSynchronyPreliminaryVersion} were proposed, devising new
SMR algorithms is still complex enough to justify a publication in a top conference
\cite{MoraruAndersenKaminsky13ThereIsMoreConsensusEgalitarianParliaments,OngaroOusterhout14SearchUnderstandableConsensusAlgorithm}.

@@

We propose a framework to make existing Generalized Consensus algorithms modular
and to enable adding new optimizations incrementally, without modification to
the existing. Generalized consensus, abbreviated GC, is a precise formulation of
State-Machine Replication that subsumes consensus. An algorithm in our framework
is composed of independent modules that can run, abort their execution, and pass
the baton to any other module on the fly.

@@

The key novelty of our framework lies in its practical applicability:
\begin{itemize}
  \item Our framework is based on Generalized Consensus, a generalization of the consensus problem 
    that allows optimizing the execution of commuting commands.
  \item Many SMR algorithms are variations on a basic structure consisting of one of two kinds of 
    quorum systems \cite{GuerraouiVukolic10RefinedQuorumSystems}. 
    The first kind, that we call resilient quorums, are used in leader-based algorithms, while the second
    kind of quorums, that we call fast quorums, are used in leader-less algorithms and can order commuting commands without synchronization.
    We provide two specifications, at an intermediate level of abstraction, that capture the essence of composing algorithms using these two kinds of quorums.
    Our two specifications allow to easily check that new modules based on resilient or fast quorums are correct, covering a large spectrum of practical algorithms. 
\end{itemize}

\begin{comment}
Allowing the execution of commuting commands to proceed without
synchronization is an important practical optimization, as commands may
often commute in practice, for example when operating on disjoint parts
of a data structure. To enable the synchronization-free execution of
commuting commands, we have based our framework on Generalized Consensus
\cite{Lamport05GeneralizeConsensus}. Traditional SMR implementations use a
separate instance of a consensus algorithm to agree on the command to execute
at each index of the command sequence on which the replicas must agree. Using
separate consensus instances makes it hard to optimize the execution of
commuting command because commutativity is not a property of a single command.
Instead of restricting agreement to a single value, the Generalized Consensus
problem requires that set of replica servers agree on the state of a growing
data structure called a command-structure set. A command-structure set is
a data structure with an append operation such that the order in which two
commuting commands are appended does not change the resulting state. Thanks to
this property, two replicas executing commuting commands in a different order
will not end up in a different state and can continue executing without further
synchronization. In message-passing systems, this property can be leveraged to
reduce the latency of processing one command to two message delays instead of
three. 
%As we will see, fast execution of commutative requests also enables the processing of read-only commands
%in two message delays, although read-only commands do not commute with write
%commands.

In practice, a SMR algorithm must support fast reconfiguration to replace
faulty components, without which the long-term stability of a system is
inevitably compromised. Our framework abstracts over how the reconfiguration
decisions are made, providing an interface allowing to plug in an external
reconfiguration master implementing the reconfiguration policy of the
system. The master decides which GC module to run next and which replica
servers will participate, removing failed replicas and adding new ones. The
reconfiguration master enables tolerating $k$ failures with $k+1$ replicas
\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication}, an important
improvement over the traditional $2k+1$ needed replicas.
\end{comment}

@@

Moreover, our interface for rounds enables a crucial optimization seen in
practical systems: introducing an external reconfiguration master which manages
the addition of new replicas and the removal of failed ones. Interestingly, as
explained in \cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication},
in a real system it is often possible to implement a reliable reconfiguration
master which allows to tolerate $f$ faults with $f+1$ replicas instead of the
traditional $2f+1$ replicas. The reconfiguration master can also make the choice
of the new type of round to execute, providing the system's reconfiguration
policy.

>>>>

The State-Machine Replication technique (abbreviated SMR) allows building
reliable services on top of unreliable hardware. The basic idea is to
replicate a service over several servers and enforce through a distributed protocol that each execute the same sequence of
deterministic commands. In such a setup, any replica can be used to determine
the result of a command. Therefore, if a replica crashes, clients can be
transparently redirected to another one and will have the illusion of accessing
a centralized service at all times.

Making SMR algorithms efficient has been a long-standing challenge
and a multitude of algorithms have been proposed. Examples include
Paxos~\cite{lamport2001paxos}, Fast Paxos~\cite{Lamport06FastPaxos},
Disk Paxos~\cite{GafniLamport03DiskPaxos}, Generalized Paxos~\cite{Lamport05GeneralizeConsensus}, FGGC~\cite{SutraShapiro11FastGenuineGeneralizedConsensus}, Chain Replication~\cite{RenesseSchneider04ChainReplicationSupportingHighThroughputAvailability},
Ring Paxos~\cite{MarandiETAL10RingPaxosHighthroughputAtomicBroadcastProtocol},
Egalitarian Paxos
\cite{MoraruAndersenKaminsky13ThereIsMoreConsensusEgalitarianParliaments},
Multi-Coordinated Paxos~\cite{CamargosSchmidtPedone07MulticoordinatedPaxos},
Cheap Paxos~\cite{LamportMassa04CheapPaxos}, Vertical Paxos
\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication}, Paxos-MIC
\cite{HurfinMoiseNarzul11AdaptiveFastPaxosMakingQuickEverlasting}, Mencius
\cite{MaoJunqueiraMarzullo08MenciusBuildingEfficientReplicatedStateMachine},
Fast Mencius~\cite{WeiETAL13FastMenciusMenciusLowCommitLatency},
the Chandra-Toueg algorithm
\cite{ChandraToueg96UnreliableFailureDetectorsReliableDistributedSystems}, the
Ben-Or algorithm~\cite{BenOr83AnotherAdvantageFreeChoiceCompletelyAsynchronous},
and Raft~\cite{OngaroOusterhout14SearchUnderstandableConsensusAlgorithm}.

Improvements in the performance and
efficiency of SMR algorithms have the potential to impact most online services, which widely rely on cloud computing:
advanced SMR algorithms can be implemented by cloud computing service providers
and benefit all their customer's applications.

However, devising new SMR algorithms, optimized with
particular situations or metrics in mind, is hard: today,
more than two decades after the first SMR algorithms~\cite{Lamport98ParttimeParliament,BirmanJoseph87ReliableCommunicationPresenceFailures,OkiLiskov88ViewstampedReplicationGeneralPrimaryCopy,DworkLynchStockmeyer84ConsensusPresencePartialSynchronyPreliminaryVersion} were proposed, devising new SMR algorithms is still complex enough to justify a publication in a top conference~\cite{MoraruAndersenKaminsky13ThereIsMoreConsensusEgalitarianParliaments,OngaroOusterhout14SearchUnderstandableConsensusAlgorithm}.

@@

We propose a framework to make existing Generalized Consensus (TODO Generalized Consensus has not been introduced) algorithms modular
and to enable adding new optimizations incrementally, without modification to
the existing. Generalized consensus, abbreviated GC, is a precise formulation of
State-Machine Replication that subsumes consensus. An algorithm in our framework
is composed of independent modules that can run, abort their execution, and pass
the baton to any other module on the fly.

@@

The key novelty of our framework lies in its practical applicability:
\begin{itemize}
  \item Our framework is based on Generalized Consensus, a generalization of the consensus problem 
    that allows optimizing the execution of commuting commands.
  \item Many SMR algorithms are variations on a basic structure consisting of one of two kinds of 
    quorum systems~\cite{GuerraouiVukolic10RefinedQuorumSystems}. 
    The first kind, that we call resilient quorums, are used in leader-based algorithms, while the second
    kind of quorums, that we call fast quorums, are used in leader-less algorithms and can order commuting commands without synchronization.
    We provide two specifications, at an intermediate level of abstraction, that capture the essence of composing algorithms using these two kinds of quorums.
    Our two specifications allow to easily check that new modules based on resilient or fast quorums are correct, covering a large spectrum of practical algorithms. 
\end{itemize}

\begin{comment}
Allowing the execution of commuting commands to proceed without
synchronization is an important practical optimization, as commands may
often commute in practice, for example when operating on disjoint parts
of a data structure. To enable the synchronization-free execution of
commuting commands, we have based our framework on Generalized Consensus~\cite{Lamport05GeneralizeConsensus}. Traditional SMR implementations use a
separate instance of a consensus algorithm to agree on the command to execute
at each index of the command sequence on which the replicas must agree. Using
separate consensus instances makes it hard to optimize the execution of
commuting command because commutativity is not a property of a single command.
Instead of restricting agreement to a single value, the Generalized Consensus
problem requires that set of replica servers agree on the state of a growing
data structure called a command-structure set. A command-structure set is
a data structure with an append operation such that the order in which two
commuting commands are appended does not change the resulting state. Thanks to
this property, two replicas executing commuting commands in a different order
will not end up in a different state and can continue executing without further
synchronization. In message-passing systems, this property can be leveraged to
reduce the latency of processing one command to two message delays instead of
three. 
%As we will see, fast execution of commutative requests also enables the processing of read-only commands
%in two message delays, although read-only commands do not commute with write
%commands.

In practice, a SMR algorithm must support fast reconfiguration to replace
faulty components, without which the long-term stability of a system is
inevitably compromised. Our framework abstracts over how the reconfiguration
decisions are made, providing an interface allowing to plug in an external
reconfiguration master implementing the reconfiguration policy of the
system. The master decides which GC module to run next and which replica
servers will participate, removing failed replicas and adding new ones. The
reconfiguration master enables tolerating $k$ failures with $k+1$ replicas~\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication}, an important
improvement over the traditional $2k+1$ needed replicas.
\end{comment}

@@

Moreover, our interface for rounds enables a crucial optimization seen in
practical systems: introducing an external reconfiguration master which manages
the addition of new replicas and the removal of failed ones. Interestingly, as
explained in~\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication},
in a real system it is often possible to implement a reliable reconfiguration
master which allows to tolerate $f$ faults with $f+1$ replicas instead of the
traditional $2f+1$ replicas. The reconfiguration master can also make the choice
of the new type of round to execute, providing the system's reconfiguration
policy.
